MRD_final_microbiome

### 2019-06-19 GlobalMicrobiome_NS.Rmd
updated

# checks that it has worked.
get_variable(pseq, "mnd")

@@

## pseq summary
```{r pseq}

pseq

```{r List of genera in phylum}

# List of Genera in the Bacteroideted Phylum
taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Bacteroidetes
print(taxa)

@@

# Summarise contents of phyloseq
summarize_phyloseq(pseq)
# Sparsity is how populated is the data with zeros

kable(head(tax_table(pseq)))
tax_table(pseq)

@@

# Prepare data for vizualisation

# Alpha diversity again
png("alpha.png")
tab <- diversities(pseq, index = "all")
stargazer(tab, type = "text",summary = NULL)
kable(head(tab))
plot(tab)
dev.off()

@@

p.low <- subset_samples(pseq, mnd == "LOW")
p.low.d <- divergence(p.low)

p.med <- subset_samples(pseq, mnd == "MEDIUM")
p.med.d <- divergence(p.med)

p.high <- subset_samples(pseq, mnd == "HIGH")
p.high.d <- divergence(p.high)

p <- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))
plot(p)

```
 The low group tends to have smaller values, indicating that the samples are more similar to the group mean, and is less heterogeneous (has smaller spread / is more homogeneous):

@@

p2 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, nrow = 2) + labs(color = "PD Prevalence")
ggsave("NMDS_facetpd.png")

p3 = plot_ordination(pseq.rel, pseq.ord, type = "sample", color = "pd")
plot(p3 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + labs(color = "PD Prevalence"))
ggsave("NMDStpd.png")

p3 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, nrow = 2) + labs(color = "PD Prevalence")
ggsave("NMDS_facetpd.png")

@@

p3 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~pd, 3) + labs(color = "AD Prevalence")
ggsave("NMDS_facetAD.png")

@@

tsne_res <- tsne_phyloseq(pseq, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
tsne_p <- plot_tsne_phyloseq(pseq, tsne_res,
  color = 'mnd', title='t-SNE (Bray)')
tsne_p + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence")
ggsave("t_sne.png")

@@

# Convert to compositional data
pseqLvH <- merge_phyloseq(p.low,p.high)
pseqLvM <- merge_phyloseq(p.low,p.med)

pseq.rel <- microbiome::transform(pseqLvH, "compositional")
otu <- abundances(pseq.rel)
meta <- meta(pseq.rel)

>>>>

# checks that it has worked.
#get_variable(pseq, "mnd")

@@

## pseq summary
```{r pseq}
pseq
```

```{r List of genera in phylum}
# List of Genera in the Bacteroideted Phylum
taxa <- map_levels(NULL, "Phylum", "Genus", pseq)$Bacteroidetes
print(taxa)
```

@@

# Summarise contents of phyloseq
summarize_phyloseq(pseq)
# Sparsity is how populated is the data with zeros
kable(head(tax_table(pseq)))
tax_table(pseq)

@@

# Prepare data for vizualisation
# Alpha diversity for the entire data set
png("alpha.png")
tab <- diversities(pseq, index = "all")
stargazer(tab, type = "text",summary = NULL)
kable(head(tab))
plot(tab)
dev.off()

@@

 The low group tends to have smaller values, indicating that the samples are more similar to the group mean, and is less heterogeneous (has smaller spread / is more homogeneous):

@@

p2 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~pd, nrow = 2) + labs(color = "PD Prevalence")
ggsave("NMDS_facetpd.png")

p3 = plot_ordination(pseq.rel, pseq.ord, type = "sample", color = "ad")
plot(p3 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + labs(color = "AD Prevalence"))
ggsave("NMDStad.png")

p3 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~ad, nrow = 2) + labs(color = "AD Prevalence")
ggsave("NMDS_facetad.png")

@@

p3 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~ad, 3) + labs(color = "AD Prevalence")
ggsave("NMDS_facetAD.png")

@@

tsne_res <- tsne_phyloseq(pseq, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
#MND
tsne_p <- plot_tsne_phyloseq(pseq, tsne_res,
  color = 'mnd', title='t-SNE (Bray)')
tsne_p + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence")
ggsave("mnd_t_sne.png")
# PD
tsne_p <- plot_tsne_phyloseq(pseq, tsne_res,
  color = 'pd', title='t-SNE (Bray)')
tsne_p + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "PD Prevalence")
ggsave("pd_t_sne.png")

# AD
tsne_p <- plot_tsne_phyloseq(pseq, tsne_res,
  color = 'ad', title='t-SNE (Bray)')
tsne_p + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "AD Prevalence")
ggsave("ad_t_sne.png")

#MND
mnd.p.low <- subset_samples(pseq, mnd == "LOW")
mnd.p.low.d <- divergence(mnd.p.low)

mnd.p.med <- subset_samples(pseq, mnd == "MEDIUM")
p.med.d <- divergence(mnd.p.med)

mnd.p.high <- subset_samples(pseq, mnd == "HIGH")
p.high.d <- divergence(mnd.p.high)

mnd.p <- boxplot(list(Low = mnd.p.low.d, Medium = mnd.p.med.d, High = mnd.p.high.d))
plot(mnd.p)

# PD

pd.p.low <- subset_samples(pseq, pd == "LOW")
pd.p.low.d <- divergence(pd.p.low)

pd.p.med <- subset_samples(pseq, pd == "MEDIUM")
p.med.d <- divergence(pd.p.med)

pd.p.high <- subset_samples(pseq, pd == "HIGH")
p.high.d <- divergence(pd.p.high)

pd.p <- boxplot(list(Low = pd.p.low.d, Medium = pd.p.med.d, High = pd.p.high.d))
plot(pd.p)

# AD

ad.p.low <- subset_samples(pseq, ad == "LOW")
ad.p.low.d <- divergence(ad.p.low)

ad.p.med <- subset_samples(pseq, ad == "MEDIUM")
p.med.d <- divergence(ad.p.med)

ad.p.high <- subset_samples(pseq, ad == "HIGH")
p.high.d <- divergence(ad.p.high)

ad.p <- boxplot(list(Low = ad.p.low.d, Medium = ad.p.med.d, High = ad.p.high.d))
plot(ad.p)

@@

# Convert to compositional data
# MND
mnd.pseqLvH <- merge_phyloseq(mnd.p.low,mnd.p.high)
mnd.pseqLvM <- merge_phyloseq(mnd.p.low,mnd.p.med)

mnd.pseq.rel <- microbiome::transform(mnd.pseqLvH, "compositional")
mnd.otu <- abundances(mnd.pseq.rel)
mnd.meta <- meta(mnd.pseq.rel)

# PD

pd.pseqLvH <- merge_phyloseq(pd.p.low,pd.p.high)
pd.pseqLvM <- merge_phyloseq(pd.p.low,pd.p.med)

pd.pseq.rel <- microbiome::transform(pd.pseqLvH, "compositional")
pd.otu <- abundances(pd.pseq.rel)
pd.meta <- meta(pd.pseq.rel)

# AD

ad.pseqLvH <- merge_phyloseq(ad.p.low,ad.p.high)
ad.pseqLvM <- merge_phyloseq(ad.p.low,ad.p.med)

ad.pseq.rel <- microbiome::transform(ad.pseqLvH, "compositional")
ad.otu <- abundances(ad.pseq.rel)
ad.meta <- meta(ad.pseq.rel)
